---
title: "SQMB Formative Assessment 1"
author: "B160039"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r attach-pkgs}
# Add the necessary packages here
library(tidyverse)
knitr::opts_knit$set(root.dir = here::here())
library(brms)
```


# Instructions

This is a guided assessment that walks your through a basic data exploration pipeline with the main aim of producing a set of relevant plots.

Follow this template to complete the assignment. Be careful with your coding style and formatting.

The data for this assessment are taken from this study: https://ling.auf.net/lingbuzz/006708.
The study looked at the "locative impersonal", i.e. *everywhere* in sentences like:

1. Is eating fish and chips a regional thing? No, these days *everywhere* in Britain eats fish and chips!
2. When Chinese food first reached Europe, people used a knife and fork, but *everywhere* in Europe eats Chinese food using chopsticks these days.

The study collected acceptability ratings of those type of sentences, using an acceptability scale from 1 to 7 (a Likert scale).

Skim through Sections 1-3 of the paper linked above for an overview of the study.

In this assignment, you will have to make some plots of a selection of the data from the study.

# Import data

First import the data. Remember to attach the necessary packages in the code chunk above.

```{r read-data}
results <- read_csv("/Users/hana/Desktop/sqmb-f1-main/data/uk_formatted_results_cleaned.csv")
```

The data frame contains 25 columns. These are the main ones:

- `Value` in the data contains the acceptability ratings.
- `Restrictor` specifies if the sentence contained a restrictor for the impersonal locative.
  - `0` non-restricted: *Everywhere eats fish and chips!*
  - `1` restricted: *Everywhere in Britain eats fish and chips!*
- `Condition` indicates which further syntactic condition the specific trial was from.
  - `gold-good` and `gold-bad` are control sentences that are expected to elicit high and low acceptability ratings respectively.
  - The other levels of `Condition` are the experimental conditions.

Mutate the values of the `Restrictor` column so that `0` becomes `non_restricted` and `1` becomes `restricted`. Can you think why this is better than using `0` and `1`?

```{r}
results <- results %>%
  mutate(
    Restrictor = ifelse (
      Restrictor %in% "0",
      "non_restricted",
      "restricted"
    )
  )
```

*==> It is more easily accessible. If someone who has not been introduced to the data beforehand opens the file, this way they will know what the values stand for. / Also if the experimenter themselves wants to return to the data much later, this way they'll know what they meant based on these labels.*

# Summarise data

Now calculate summaries measures for `Value`, `Restrictor` and `Condition` and any other column you like. Also, count occurrences of `Value` across `Restrictor` and `Condition`, separately and combined.

# Summarise measures for Value

==> The most frequent value (i.e., the mode) for *Value* is 7 (= 1610). (I chose the mode as the summary measure of central tendency for Value, Restrictor as well as Condition because all of these values are categorical discrete therefore mean and median would be inappropriate for them.)

```{r}
count(results, Value)
```

# Summarise measures for Restrictor

==> The 2 conditions of the *Restrictor* receive the same total of values (= 2645) and thus do not have a mode.

```{r}
count(results, Restrictor)
```

# Summarise measures for Condition

=> The most common value (i.e., the mode) for *Condition* is *modal* (= 920).

```{r}
count(results, Condition)
```

# Count occurrences of `Value` across `Restrictor` and `Condition`, separately and combined

```{r}
count (results, Value, Restrictor)
```


```{r}
count (results, Value, Condition)
```

```{r}
count (results, Value, Restrictor, Condition)
```


# Plot

Produce the following plots. For each, write a brief description of the plot and patterns you notice in the plot.

## Plot 1

Create a simple plot which shows the total *counts* of each value in `Value`.

```{r}
ggplot(
  data = results,
  mapping = aes(x = Value)
) +
  geom_bar() +
  labs (title = "Counts for each value of acceptability rates (1-7)")
```

Describe the plot.

==> This bar chart displays the counts of times each *Value* from 1-7 was chosen as the participants' response, e.g., Value *2* was chosen approximately 450 times. What this chart shows us is that it was the 'extreme' values (i.e., 1 and 7) that received more attention - participants thus aimed for very binary response ratings. The highest count received the Value *7*, while the lowest got Value *3*.

## Plot 2

Create a plot that shows the counts for each `Value` depending on `Restrictor`.

```{r}
results %>%
  ggplot(aes(Value, fill = Restrictor)) +
  geom_bar() +
  facet_grid (cols = vars(Restrictor)) +
  labs (title = "Acceptability rates (1-7) in 2 conditions of Restricted/Non-restricted")
```

Describe the plot.

==> This bar chart displays the counts for *Values* 1-7 given 2 conditions of *Restrictor* - the *non-restricted* condition where the sentence did not contain a restrictor, e.g., *Everywhere eats fish and chips!*, and the *restricted* condition where the sentence did contain a restrictor, e.g., *Everywhere in Britain eats fish and chips!* This plot shows us that the *restricted* condition got selected more often in the higher-range values such as 5, 6, 7; in other words was rated as more acceptable. In contrast, the *non-restricted* condition was preferably selected in the lower-range values such as 1 and 2, and thus was deemed more often as less acceptable.

## Plot 3

Now add `Condition` to the mix with `Values` and `Restrictor`.
You can pick any specific solution (colour/fill, faceting, etc...).

```{r}
results %>%
  ggplot(aes(Value, fill = Condition)) +
  geom_bar() +
  facet_grid (cols = vars(Restrictor)) +
  labs (title = "Acceptability rates with stacked Conditions divided for 2 qualities of Restrictor")
```

Describe the plot.

==> Similarly to Plot 2, also Plot 3 displays the counts of *Value* 1-7 given the 2 conditions of *Restrictor*, but this time in combination with *Condition* that indicates which further syntactic condition the specific trial was from (*gold-good* and *gold-bad* are control sentences that are expected to elicit high and low acceptability ratings respectively). While the value columns for the 2 qualities of *Restrictor* remain the same as in Plot 2, Plot 3 offers us better insight into the respective distribution of the various *Conditions*. The *adverb*, *binding* and *modal* Conditions seemed to get the highest ratings of acceptability. In contrast, the *gold-good* Condition was rated more often as least acceptable - a finding contrasting the expressed expectations.

## Plot 4

Now create a last plot that includes any of the remaining columns in the data.

```{r}
results %>%
  ggplot(aes(Value, fill = Condition)) +
  geom_bar() +
  facet_grid (cols = vars(Restrictor), rows = vars(Gender)) +
  labs (title = "Acceptability rates with stacked Conditions divided for Restrictor and Gender")
```

Describe the plot.

==> Plot 4 then showcases that while there were discrepancies between the total of *Male* and *Female* participants (size of columns), the two categories of participants did not differ significantly in their ratings. Acceptability ratings of the "locative impersonal" are thus not dependent on the gender of the respective 'rators'.